
---
title: "Austin-Bergstrom Interational Airport - Visualization"
output: github_document
---



```{r}
abia = read.csv('ABIA.csv', header=TRUE)
abia = abia[,-1]
abia$ElapsedDelay = abia$CRSElapsedTime - abia$ActualElapsedTime
abia$TotalDelay = abia$ElapsedDelay + abia$ArrDelay + abia$DepDelay
abia$depart = ifelse(abia$Origin=='AUS','Depart','Arrival')
abia$early = ifelse(abia$TotalDelay<0,abia$TotalDelay,0)
abia$delayed = ifelse(abia$TotalDelay>0,1,0)
abia$Week = ifelse(abia$DayOfWeek<5,"Weekday","Weekend")
```


## 1) Delay by Airways

```{r}
library(ggplot2)
prop = xtabs(~delayed + UniqueCarrier, data=abia) 
prop = prop.table(prop, margin=2)
prop = as.data.frame(prop)
prop = prop[prop$delayed==1,]
prop$UniqueCarrier <- factor(prop$UniqueCarrier, levels = prop$UniqueCarrier[order(-prop$Freq)])
ggplot(prop, aes(x=UniqueCarrier,y=Freq))+
  geom_bar(stat="identity", fill="#FF9999", colour="black")+
  geom_text(aes(label=round(Freq,2)), vjust=1.6, color="black",size=3.5)+
  ylab("Proportion of delayed flight")
```

Atlantic Southeast Airlines(EV) has the highest delay rate with 53% followed by Southwest Airlines(WN) with 52%. US Airways(US) has the lowest delay rate with 22%.

```{r}
prop2 = xtabs(TotalDelay~UniqueCarrier , aggregate(TotalDelay~UniqueCarrier,abia,mean))
prop2 = as.data.frame(prop2)
prop2$UniqueCarrier <- factor(prop2$UniqueCarrier, levels = prop2$UniqueCarrier[order(-prop2$Freq)])
ggplot(prop2, aes(x=UniqueCarrier,y=Freq))+
  geom_bar(stat="identity", fill="#009999", colour="black")+
  geom_text(aes(label=round(Freq,2)), vjust=1.6, color="black",size=3.5)+
  ylab("Average delayed time")
```

Atlantic Southeast Airlines(EV) has also the highest average delay time with 33 minutes followed by Comair(oh) Airlines with 25 minutes. US Airways has the lowest average delay rate with 2 minutes.


## 2) Delay by month

```{r}
prop = xtabs(~delayed + Month, data=abia)
prop = prop.table(prop,margin=2)
prop = as.data.frame(prop)
prop = prop[prop$delayed==1,]
prop$Month <- factor(prop$Month, levels = prop$Month[order(-prop$Freq)])
ggplot(prop, aes(x=Month,y=Freq))+
  geom_bar(stat="identity", fill="#FF9999", colour="black")+
  geom_text(aes(label=round(Freq,2)), vjust=1.6, color="black",size=3.5)+
  ylab("Proportion of delayed flight")
```

Months in vacation, which is December, March, June, have higher proportion of delayed flight. However, fall season from September to November, the rate tend to be lower.

```{r}
prop = xtabs(TotalDelay~Month , aggregate(TotalDelay~Month,abia,mean))
prop = as.data.frame(prop)
prop$Month <- factor(prop$Month, levels = prop$Month[order(-prop$Freq)])
ggplot(prop, aes(x=Month,y=Freq)) + geom_bar(stat="identity")+
  geom_bar(stat="identity", fill="#009999", colour="black")+
  geom_text(aes(label=round(Freq,2)), vjust=1.6, color="black",size=3.5)+
  ylab("Average delayed time")
```

Similar to the proportion of delay flight, months in vacation, which is December, March, June, have higher delayed time. However, fall season from September to November, the delay time tend to be lower.

## 3) Delay by airport(Top 10)

```{r}
abia_depart = abia[which(abia$depart == 'Depart'),]
prop = xtabs(~delayed + Dest, data=abia_depart)
prop = prop.table(prop,margin=2)
prop = as.data.frame(prop)
prop = prop[prop$delayed==1,]
prop$Dest <- factor(prop$Dest, levels = prop$Dest[order(-prop$Freq)])

ggplot(na.omit(prop[1:10,]), aes(x=Dest,y=Freq))+
  geom_bar(stat="identity", fill="#FF9999", colour="black")+
  geom_text(aes(label=round(Freq,2)), vjust=1.6, color="black",size=3.5)+
  ylab("Proportion of delayed flight")+
  xlab("Destination")
```

Flight from Austin to Balitimore/Washington Airport has the highest delay rate with 53%. Nashville International Airport and Dallas Airport also have delay rate over 40%.


```{r}
prop = xtabs(TotalDelay~Dest, aggregate(TotalDelay~Dest,abia_depart,mean))
prop = as.data.frame(prop)
prop$Dest <- factor(prop$Dest, levels = prop$Dest[order(-prop$Freq)])
prop = prop[which(prop$Freq>0),]
ggplot(prop[1:10,], aes(x=Dest,y=Freq)) + geom_bar(stat="identity")+
  geom_bar(stat="identity", fill="#009999", colour="black")+
  geom_text(aes(label=round(Freq,2)), vjust=1.6, color="black",size=3.5)+
  ylab("Average delayed time")+
  xlab("Destination")
```

Flight from Austin to Atlanta has the longest average delay time with 21 minutes, and all other routes has less than 20 minutes of average delay time.

## 4) Delay by DepartTime and Weekday/Weekend

```{r}
ggplot(data = abia) + 
  geom_point(mapping = aes(x = DepTime, y = TotalDelay, color = Week))

```

Flights departing late night tend to be delayed more. Also, flight on weekend tent to be delayed more than weekdays.



## 5) Early arrival by Airways

```{r}
ggplot(data = abia_depart) + 
  geom_point(mapping = aes(x = UniqueCarrier, y = early), color='steelblue')
```

